Comparing human and model-based forecasts of COVID-19 in Germany and Poland

### 2022-09-14 manuscript/manuscript.tex
attempt to update the way references are cited

\cite{mcgowanCollaborativeEffortsForecast2019}

Infectious disease modelling has a long tradition and has helped inform
public health decisions both through scenario modelling, as well as
actual forecasts of (among others) influenza {[}e.g. 1,2--4{]}, dengue
fever {[}e.g. 5,6,7{]}, ebola {[}e.g. 8,9{]}, chikungunya {[}e.g.
10,11{]} and now COVID-19 {[}e.g. 12,13--17{]}. Applications of
epidemiological models differ in the way they make statements about the
future. Forecasts aim to predict the future as it will occur, while
scenario modelling and projections aim to represent what the future
could look like under certain scenario assumptions or if conditions
stayed the same as they were in the past. Forecasts can be judged by
comparing them against observed data. Since it is much harder to fairly
assess the accuracy and usefulness of projections and scenario modelling
in the same way, this work focuses on forecasts, which represent only a
subset of all epidemiological modelling.

Since March 2020, forecasts of COVID-19 from multiple teams have been
collected, aggregated and compared by Forecast Hubs such as the US
Forecast Hub {[}13,14{]}, the German and Polish Forecast Hub {[}15,16{]}
and the European Forecast Hub {[}17{]}. Often, different individual
forecasts are combined into a single forecast, e.g.~by taking the mean
or median of all forecasts. These ensemble forecasts usually tend to
perform better and more consistently than individual forecasts (see e.g.
{[}6{]}; {[}18{]}).

@@

Previous work has examined such direct human forecasts in various
contexts, such as geopolitics {[}19,20{]}, meta-science {[}21,22{]},
sports {[}23{]} and epidemiology {[}11,24,25{]}. Several prediction
platforms {[}26--28{]} and prediction markets {[}29{]} have been created
to collate expert and non-expert predictions. However, with the notable
exception of {[}11{]}, these forecasts were not designed to be evaluated
alongside model-based forecasts and usually follow their own (often
binary) prediction formats. Direct human forecasts may be able to take
into account insights and relationships between variables which are hard
to specify using epidemiological models. However, it is not entirely
clear in which situations human forecasts perform well or badly. For
example, {[}11{]} found that humans could outperform computer models at
predicting the 2014/15 and 2015/16 flu season in the US, a setting where
the disease was well known and information about previous seasons was
available. However, humans tended to do slightly worse at predicting the
2014/15 outbreak of chikungunya in the Americas, a disease previously
largely unobserved and unknown in these regions at the time.

In this study, we analyse the performance of direct human forecasts
relative to model-based forecasts and discuss the added benefit of
epidemiological modelling over human judgement alone. As a case study,
we use different forecasts, involving varying degrees of human
intervention, which we submitted in real time to the German and Polish
Forecast Hub. In contrast to {[}11{]} we elicited not only point
predictions, but full predictive distributions (``probabilistic
forecasts'', see e.g. {[}30{]}) from participants. This allows us to
compare not only predictive accuracy, but also how well human
forecasters and model-based forecasts were able to quantify forecast
uncertainty.

@@

We created and submitted the following forecasts to the German and
Polish Forecast Hub: 1) a direct human forecast (henceforth called
``crowd forecast''), elicited from participants through a web
application {[}31{]} and 2) two semi-mechanistic model-based forecasts
(``renewal model'' and ``convolution model'') informed by basic
assumptions about COVID-19 epidemiology. While the two semi-mechanistic
forecasts were necessarily shaped by our implicit assumptions and
decisions, they were designed such as to minimise the amount of human
intervention involved. For example, we refrained from adjusting model
outputs or refining the models based on past performance. Forecasts were
created in real time over a period of 21 weeks from October 12th 2020
until March 1st 2021 and submitted to the German and Polish Forecast hub
{[}15,16{]}. All code and tools necessary to generate the forecasts and
make a forecast submission are available in the
\texttt{covid.german.forecasts} R package {[}32{]}. This repository also
contains a record of all forecasts submitted to the German and Polish
Forecast Hub. Forecasts were evaluated using a variety of scoring
metrics and compared among each other and against an ensemble of all
other models submitted to the German and Polish Forecast Hub.

@@

The German and Polish Forecast Hub (now mostly merged into the {[}17{]})
elicits predictions for various COVID-19 related forecast targets from
different research groups every week. Forecasts had to be made every
Monday (with submissions allowed until Tuesday 3pm) and were permitted
to use any data that was available by Monday 11.59pm. We submitted
forecasts for incident and cumulative weekly reported numbers of cases
of and deaths from COVID-19 on a national level in Germany and Poland
over a one to four week forecast horizon. Forecasts were submitted on
Mondays, but weeks were defined as ending on a Saturday (and starting on
Sunday), meaning that forecast horizons were in fact 5, 12, 19 and 26
days. Submissions were required in a quantile-based format with 23
quantiles of each output measure at levels 0.01, 0.025, 0.05, 0.10,
0.15, \ldots, 0.95, 0.975, 0.99. Forecasts submitted to the Forecast Hub
were combined into different ensembles every week, with the median
ensemble (i.e., the \(\alpha\)-quantile of the ensemble is given by the
median of all submitted \(\alpha\)-quantiles) being the default ensemble
shown on all official Forecast hub visualisations
(https://kitmetricslab.github.io/forecasthub/forecast).

Data on daily reported test positive cases and deaths linked to COVID-19
were provided by the organisers of the German and Polish Forecast hub.
Until December 14th, 2020, these data were sourced from the European
Centre for Disease Control {[}33{]}. After ECDC stopped publishing daily
data, observations were sourced from the Robert Koch Institute (RKI) and
the Polish Ministry of Health for the remainder of the submission period
{[}34{]}. These data are subject to reporting artefacts, (such as for
example delayed case reporting in Poland on the 24th November,
{[}35{]}), changes in reporting over time, and variation in testing
regimes (for example in Germany from the 11th of November on, {[}36{]}).
The ECDC data as well as the data published by the Polish Ministry of
Health were also subject to data revisions, although most of them (with
a notable exception of a data update for October 12 2020 in Germany)
only affected daily, not weekly data (see
\nameref{fig:daily-truth-update} and \nameref{fig:weekly-truth-update}
Figs).

@@

Our crowd forecasts were created as an ensemble of forecasts made by
individual participants every week through a web application
(https://cmmid-lshtm.shinyapps.io/crowd-forecast/). Weekly forecasts had
to be submitted before Tuesday 12pm every week, but participants were
asked to only use any information or data that was already available by
Monday night. The application was built using the \texttt{shiny} and
\texttt{golem} R packages {[}37,38{]} and is available in the
\texttt{crowdforecastr} R package {[}31{]}. To make a forecast in the
application participants could select a predictive distribution (with
the default being log-normal) to represent the probability that the
forecasted quantity took certain values. Median and width of the
uncertainty could be adjusted by either interacting with a figure
showing their forecast or providing numerical values (see screenshot in
\nameref{fig:screenshot}). The default shown was a repetition of the
last known observation with constant uncertainty around it computed as
the standard deviation of the last four changes in weekly log observed
forecasts (i.e.~as
\(\sigma(log(value4) - log(value3), log(value3) - log(value2), \ldots )\)).
A comparison of the crowd forecasts against the default baseline shown
in the application is displayed in \nameref{fig:compare-forecasters}.
Our interface also allowed participants to view past observations based
on the hub data, as well as their forecasts, on a logarithmic scale and
presented additional contextual COVID-19 data sourced from {[}39{]}.
These data included, for example, notifications of both test positive
COVID-19 cases and COVID-19 linked deaths and the number of COVID-19
tests conducted over time. From November 26 2020 on we displayed weekly
small reports with a visualisation of past forecasts and scores on our
website, epiforecasts.io.

@@

We used two Bayesian semi-mechanistic models from the \texttt{EpiNow2} R
package (version 1.3.3) as our model-based forecasts {[}40{]}. The first
of these models, here called ``renewal model'', used the renewal
equation {[}41{]} to predict reported cases and deaths (see details in
\nameref{txt:details-models}). It estimated the effective reproduction
number \(R_t\) (the average number of people each person infected at
time t is expected to infect in turn) and modelled future infections as
a weighted sum of past infection multiplied by \(R_t\). \(R_t\) was
assumed to stay constant beyond the forecast date, roughly corresponding
to continuing the latest exponential trend in infections. On the 9th of
November we altered the date when \(R_t\) was assumed to be constant
from two weeks prior to the date of the forecast to the forecast date,
which we found to yield a more stable \(R_t\) estimate. Reported case
and death notifications were obtained by convolving predicted infections
over data-based delay distributions {[}40,42--44{]} to model the time
between infection and report date. The renewal model was used to predict
cases as well as deaths with forecasts being generated for each target
separately. Death forecasts from the renewal model were therefore not
informed by past cases. One submission of the renewal model on December
28th 2020 was delayed and therefore not included in the official
Forecast hub ensemble.

The second model (``convolution model'', see details in
\nameref{txt:details-models}) was only used to forecast deaths and was
added later, starting December 7th 2020 (with the first forecast from
December 7th suffering from a software bug and therefore disregarded in
all further analyses). The convolution model was submitted, but never
included in the official Forecast hub ensemble due to concerns that it
could be too similar to the renewal model. The convolution model
predicted deaths as a fraction of infected people who would die with
some delay, by using a convolution of reported cases with a distribution
that described the delay from case report to death and a scaling factor
(the case-fatality ratio). Both the renewal and the convolution model
used daily observations and assumed a negative binomial observation
model with a multiplicative day-of-the-week effect {[}40{]}.

Line list data used to inform the prior for the delay from symptom onset
to test positive case report or death in the model-based forecasts was
sourced from {[}45{]} with data available up to the 1st of August. All
model fitting was done using Markov-chain Monte Carlo (MCMC) in stan
{[}46{]} with each location and forecast target being fitted separately.

@@

For the main analysis we focused mostly on two week ahead forecasts, as
COVID-19 forecasts, especially for cases, were in the past found to have
poor predictive performance beyond this horizon {[}15{]}. Forecasts for
cases were scored using the full period from October 2020 until March
2021. To ensure comparability between models, all death forecasts were
scored using only the period from December 14th on, where all models
including the convolution model were available. To ensure robustness of
our results we conducted a sensitivity analysis where all forecasts
(including cases) were scored only over the later period for which all
forecasts were available (see \nameref{fig:agg-performance-all-late} and
\nameref{tab:score-table-late-2} and \nameref{tab:score-table-late-4}s).
Results remained broadly unchanged.

Forecasts were analysed using the following scoring metrics: The
weighted interval score (WIS) {[}47{]}, the absolute error, relative
bias, and empirical coverage of the 50\% and 90\% prediction intervals.
The WIS is a proper scoring rule {[}48{]}, meaning that in expectation
the score is optimised by reporting a predictive distribution that is
identical to the true data-generating distribution. Forecasters are
therefore incentivised to report their true belief about the future. The
WIS can be understood as a generalisation of the absolute error to
quantile-based forecasts (also meaning that smaller values are better)
and can be decomposed into three separate penalties: forecast spread
(i.e.~uncertainty of forecasts), over-prediction and under-prediction.
While the over- and under-prediction components of the WIS capture the
amount of over-prediction and under-prediction in absolute terms, we
also look at a relative tendency to make biased forecasts. The bias
metric {[}9{]} we use captures how much probability mass of the forecast
was above or below the true value (mapped to values between -1 and 1)
and therefore represents a general tendency to over- or under-predict in
relative terms. A value of -1 implies that all quantiles of the
predictive distribution are below the observed value and a value of 1
that all quantiles are above the observed value. Empirical coverage is
the percentage of observed values that fall inside a given prediction
interval (e.g.~how many observed values fall inside all 50\% prediction
intervals). Scoring metrics are explained in more detail in
\nameref{tab:scoring-metrics}. All scores were calculated using the
\texttt{scoringutils} R package {[}49{]}.

@@

To varying degrees, all forecasts exhibited trend-following behaviour
and were rarely able to predict a change in trend before it had
happened. For example, all forecasts failed to predict the change in
trend from increase to decrease that happened in November in Germany and
severely overshot reported cases (Fig \ref{fig:forecasts-and-truth}A).
This was most striking for the renewal model, which extrapolated
unconstrained exponential growth based on the recent past of
observations. The Hub ensemble and the crowd forecast, which had both
been under-predicting throughout October, also failed to predict the
change in trend after cases peaked, but less severely so. Human
forecasters, possibly aware of the semi-lockdown announced on November
2nd 2020 {[}50{]} and the change in the testing regime (with stricter
test criteria) on November 11th 2020 {[}36{]}, were fastest to adapt to
the new trend, and the Hub ensemble slowest. In December, cases rose
again in Germany, with all models under-predicting this growth to
varying extents. As in October, the renewal model captured the phase of
exponential growth in cases slightly better than other approaches, but
again overshot when reported case numbers fell over Christmas. The large
variance in predictions in January in Germany (severe under-prediction
followed by severe over-prediction) may in part be caused by the fact
that the renewal model operated on daily data and therefore was
susceptible to fluctuations in daily reporting around Christmas that
would not have influenced on weekly reporting. Similar trends in
performance were evident in Poland, with the crowd forecast quickest at
adapting to the change in trend in November. In general, there were
fewer large outlier forecasts in Poland and in particular the renewal
model performed more in line with other forecasts there.

@@

Past studies have evaluated the performance of model-based forecasting
approaches as well as human experts and non-experts in various contexts.
However, most of these studies either focused only on the evaluation of
(expert-tuned) model-based approaches {[}e.g. 12,13,14{]}, or
exclusively on human forecasts {[}19,20,24,25{]}. In contrast, we
directly compared human and model-based forecasts. This is similar to
the approach taken by {[}11{]}, but extends it in several ways. While
Farrow et al.~only asked for point predictions and constructed a
predictive distribution from these, we asked participants to provide a
full predictive distribution, allowing us to compare human forecasts and
models without any further assumptions, as well as to analyse how humans
quantified their uncertainty. In addition, we compared crowd forecasts
to two semi-mechanistic models informed by basic epidemiological
knowledge of COVID-19, allowing us to assess not only relative
performance but also to analyse qualitative differences between human
judgement and model-based insight. In terms of interpretability of the
results, exact knowledge of our two models, as well as focus on a
limited set of targets and locations was a major advantage of our study
compared to larger studies conducted by the Forecast Hubs
{[}12--15,17{]}.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-mcgowanCollaborativeEffortsForecast2019}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{McGowan CJ, Biggerstaff M, Johansson M, Apfeldorf KM,
Ben-Nun M, Brooks L, et al. Collaborative efforts to forecast seasonal
influenza in the {United States}, 2015--2016. Scientific Reports.
2019;9: 683.
doi:\href{https://doi.org/10.1038/s41598-018-36361-9}{10.1038/s41598-018-36361-9}}

\leavevmode\vadjust pre{\hypertarget{ref-reichCollaborativeMultiyearMultimodel2019}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{Reich NG, Brooks LC, Fox SJ, Kandula S, McGowan CJ,
Moore E, et al. A collaborative multiyear, multimodel assessment of
seasonal influenza forecasting in the {United States}. PNAS. 2019;116:
3146--3154.
doi:\href{https://doi.org/10.1073/pnas.1812594116}{10.1073/pnas.1812594116}}

\leavevmode\vadjust pre{\hypertarget{ref-shamanForecastingSeasonalOutbreaks2012}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{Shaman J, Karspeck A. Forecasting seasonal outbreaks of
influenza. PNAS. 2012;109: 20425--20430.
doi:\href{https://doi.org/10.1073/pnas.1208772109}{10.1073/pnas.1208772109}}

\leavevmode\vadjust pre{\hypertarget{ref-biggerstaffResultsCentersDisease2016}{}}%
\CSLLeftMargin{4. }%
\CSLRightInline{Biggerstaff M, Alper D, Dredze M, Fox S, Fung IC-H,
Hickmann KS, et al. Results from the centers for disease control and
prevention's predict the 2013--2014 {Influenza Season Challenge}. BMC
Infectious Diseases. 2016;16: 357.
doi:\href{https://doi.org/10.1186/s12879-016-1669-x}{10.1186/s12879-016-1669-x}}

\leavevmode\vadjust pre{\hypertarget{ref-johanssonOpenChallengeAdvance2019}{}}%
\CSLLeftMargin{5. }%
\CSLRightInline{Johansson MA, Apfeldorf KM, Dobson S, Devita J, Buczak
AL, Baugher B, et al. An open challenge to advance probabilistic
forecasting for dengue epidemics. PNAS. 2019;116: 24268--24274.
doi:\href{https://doi.org/10.1073/pnas.1909865116}{10.1073/pnas.1909865116}}

\leavevmode\vadjust pre{\hypertarget{ref-yamanaSuperensembleForecastsDengue2016}{}}%
\CSLLeftMargin{6. }%
\CSLRightInline{Yamana TK, Kandula S, Shaman J. Superensemble forecasts
of dengue outbreaks. Journal of The Royal Society Interface. 2016;13:
20160410.
doi:\href{https://doi.org/10.1098/rsif.2016.0410}{10.1098/rsif.2016.0410}}

\leavevmode\vadjust pre{\hypertarget{ref-colon-gonzalezProbabilisticSeasonalDengue2021}{}}%
\CSLLeftMargin{7. }%
\CSLRightInline{Colón-González FJ, Bastos LS, Hofmann B, Hopkin A,
Harpham Q, Crocker T, et al. Probabilistic seasonal dengue forecasting
in {Vietnam}: {A} modelling study using superensembles. PLOS Medicine.
2021;18: e1003542.
doi:\href{https://doi.org/10.1371/journal.pmed.1003542}{10.1371/journal.pmed.1003542}}

>>>>

Infectious disease modelling has a long tradition and has helped inform
public health decisions both through scenario modelling, as well as
actual forecasts of (among others) influenza
(e.g.~\cite{mcgowanCollaborativeEffortsForecast2019, reichCollaborativeMultiyearMultimodel2019,  shamanForecastingSeasonalOutbreaks2012, biggerstaffResultsCentersDisease2016},
dengue fever {[}e.g. 1,2,3{]}, ebola {[}e.g. 4,5{]}, chikungunya {[}e.g.
6,7{]} and now COVID-19 {[}e.g. 8,9--13{]}. Applications of
epidemiological models differ in the way they make statements about the
future. Forecasts aim to predict the future as it will occur, while
scenario modelling and projections aim to represent what the future
could look like under certain scenario assumptions or if conditions
stayed the same as they were in the past. Forecasts can be judged by
comparing them against observed data. Since it is much harder to fairly
assess the accuracy and usefulness of projections and scenario modelling
in the same way, this work focuses on forecasts, which represent only a
subset of all epidemiological modelling.

Since March 2020, forecasts of COVID-19 from multiple teams have been
collected, aggregated and compared by Forecast Hubs such as the US
Forecast Hub {[}9,10{]}, the German and Polish Forecast Hub {[}11,12{]}
and the European Forecast Hub {[}13{]}. Often, different individual
forecasts are combined into a single forecast, e.g.~by taking the mean
or median of all forecasts. These ensemble forecasts usually tend to
perform better and more consistently than individual forecasts (see e.g.
{[}2{]}; {[}14{]}).

@@

Previous work has examined such direct human forecasts in various
contexts, such as geopolitics {[}15,16{]}, meta-science {[}17,18{]},
sports {[}19{]} and epidemiology {[}7,20,21{]}. Several prediction
platforms {[}22--24{]} and prediction markets {[}25{]} have been created
to collate expert and non-expert predictions. However, with the notable
exception of {[}7{]}, these forecasts were not designed to be evaluated
alongside model-based forecasts and usually follow their own (often
binary) prediction formats. Direct human forecasts may be able to take
into account insights and relationships between variables which are hard
to specify using epidemiological models. However, it is not entirely
clear in which situations human forecasts perform well or badly. For
example, {[}7{]} found that humans could outperform computer models at
predicting the 2014/15 and 2015/16 flu season in the US, a setting where
the disease was well known and information about previous seasons was
available. However, humans tended to do slightly worse at predicting the
2014/15 outbreak of chikungunya in the Americas, a disease previously
largely unobserved and unknown in these regions at the time.

In this study, we analyse the performance of direct human forecasts
relative to model-based forecasts and discuss the added benefit of
epidemiological modelling over human judgement alone. As a case study,
we use different forecasts, involving varying degrees of human
intervention, which we submitted in real time to the German and Polish
Forecast Hub. In contrast to {[}7{]} we elicited not only point
predictions, but full predictive distributions (``probabilistic
forecasts'', see e.g. {[}26{]}) from participants. This allows us to
compare not only predictive accuracy, but also how well human
forecasters and model-based forecasts were able to quantify forecast
uncertainty.

@@

We created and submitted the following forecasts to the German and
Polish Forecast Hub: 1) a direct human forecast (henceforth called
``crowd forecast''), elicited from participants through a web
application {[}27{]} and 2) two semi-mechanistic model-based forecasts
(``renewal model'' and ``convolution model'') informed by basic
assumptions about COVID-19 epidemiology. While the two semi-mechanistic
forecasts were necessarily shaped by our implicit assumptions and
decisions, they were designed such as to minimise the amount of human
intervention involved. For example, we refrained from adjusting model
outputs or refining the models based on past performance. Forecasts were
created in real time over a period of 21 weeks from October 12th 2020
until March 1st 2021 and submitted to the German and Polish Forecast hub
{[}11,12{]}. All code and tools necessary to generate the forecasts and
make a forecast submission are available in the
\texttt{covid.german.forecasts} R package {[}28{]}. This repository also
contains a record of all forecasts submitted to the German and Polish
Forecast Hub. Forecasts were evaluated using a variety of scoring
metrics and compared among each other and against an ensemble of all
other models submitted to the German and Polish Forecast Hub.

@@

The German and Polish Forecast Hub (now mostly merged into the {[}13{]})
elicits predictions for various COVID-19 related forecast targets from
different research groups every week. Forecasts had to be made every
Monday (with submissions allowed until Tuesday 3pm) and were permitted
to use any data that was available by Monday 11.59pm. We submitted
forecasts for incident and cumulative weekly reported numbers of cases
of and deaths from COVID-19 on a national level in Germany and Poland
over a one to four week forecast horizon. Forecasts were submitted on
Mondays, but weeks were defined as ending on a Saturday (and starting on
Sunday), meaning that forecast horizons were in fact 5, 12, 19 and 26
days. Submissions were required in a quantile-based format with 23
quantiles of each output measure at levels 0.01, 0.025, 0.05, 0.10,
0.15, \ldots, 0.95, 0.975, 0.99. Forecasts submitted to the Forecast Hub
were combined into different ensembles every week, with the median
ensemble (i.e., the \(\alpha\)-quantile of the ensemble is given by the
median of all submitted \(\alpha\)-quantiles) being the default ensemble
shown on all official Forecast hub visualisations
(https://kitmetricslab.github.io/forecasthub/forecast).

Data on daily reported test positive cases and deaths linked to COVID-19
were provided by the organisers of the German and Polish Forecast hub.
Until December 14th, 2020, these data were sourced from the European
Centre for Disease Control {[}29{]}. After ECDC stopped publishing daily
data, observations were sourced from the Robert Koch Institute (RKI) and
the Polish Ministry of Health for the remainder of the submission period
{[}30{]}. These data are subject to reporting artefacts, (such as for
example delayed case reporting in Poland on the 24th November,
{[}31{]}), changes in reporting over time, and variation in testing
regimes (for example in Germany from the 11th of November on, {[}32{]}).
The ECDC data as well as the data published by the Polish Ministry of
Health were also subject to data revisions, although most of them (with
a notable exception of a data update for October 12 2020 in Germany)
only affected daily, not weekly data (see
\nameref{fig:daily-truth-update} and \nameref{fig:weekly-truth-update}
Figs).

@@

Our crowd forecasts were created as an ensemble of forecasts made by
individual participants every week through a web application
(https://cmmid-lshtm.shinyapps.io/crowd-forecast/). Weekly forecasts had
to be submitted before Tuesday 12pm every week, but participants were
asked to only use any information or data that was already available by
Monday night. The application was built using the \texttt{shiny} and
\texttt{golem} R packages {[}33,34{]} and is available in the
\texttt{crowdforecastr} R package {[}27{]}. To make a forecast in the
application participants could select a predictive distribution (with
the default being log-normal) to represent the probability that the
forecasted quantity took certain values. Median and width of the
uncertainty could be adjusted by either interacting with a figure
showing their forecast or providing numerical values (see screenshot in
\nameref{fig:screenshot}). The default shown was a repetition of the
last known observation with constant uncertainty around it computed as
the standard deviation of the last four changes in weekly log observed
forecasts (i.e.~as
\(\sigma(log(value4) - log(value3), log(value3) - log(value2), \ldots )\)).
A comparison of the crowd forecasts against the default baseline shown
in the application is displayed in \nameref{fig:compare-forecasters}.
Our interface also allowed participants to view past observations based
on the hub data, as well as their forecasts, on a logarithmic scale and
presented additional contextual COVID-19 data sourced from {[}35{]}.
These data included, for example, notifications of both test positive
COVID-19 cases and COVID-19 linked deaths and the number of COVID-19
tests conducted over time. From November 26 2020 on we displayed weekly
small reports with a visualisation of past forecasts and scores on our
website, epiforecasts.io.

@@

We used two Bayesian semi-mechanistic models from the \texttt{EpiNow2} R
package (version 1.3.3) as our model-based forecasts {[}36{]}. The first
of these models, here called ``renewal model'', used the renewal
equation {[}37{]} to predict reported cases and deaths (see details in
\nameref{txt:details-models}). It estimated the effective reproduction
number \(R_t\) (the average number of people each person infected at
time t is expected to infect in turn) and modelled future infections as
a weighted sum of past infection multiplied by \(R_t\). \(R_t\) was
assumed to stay constant beyond the forecast date, roughly corresponding
to continuing the latest exponential trend in infections. On the 9th of
November we altered the date when \(R_t\) was assumed to be constant
from two weeks prior to the date of the forecast to the forecast date,
which we found to yield a more stable \(R_t\) estimate. Reported case
and death notifications were obtained by convolving predicted infections
over data-based delay distributions {[}36,38--40{]} to model the time
between infection and report date. The renewal model was used to predict
cases as well as deaths with forecasts being generated for each target
separately. Death forecasts from the renewal model were therefore not
informed by past cases. One submission of the renewal model on December
28th 2020 was delayed and therefore not included in the official
Forecast hub ensemble.

The second model (``convolution model'', see details in
\nameref{txt:details-models}) was only used to forecast deaths and was
added later, starting December 7th 2020 (with the first forecast from
December 7th suffering from a software bug and therefore disregarded in
all further analyses). The convolution model was submitted, but never
included in the official Forecast hub ensemble due to concerns that it
could be too similar to the renewal model. The convolution model
predicted deaths as a fraction of infected people who would die with
some delay, by using a convolution of reported cases with a distribution
that described the delay from case report to death and a scaling factor
(the case-fatality ratio). Both the renewal and the convolution model
used daily observations and assumed a negative binomial observation
model with a multiplicative day-of-the-week effect {[}36{]}.

Line list data used to inform the prior for the delay from symptom onset
to test positive case report or death in the model-based forecasts was
sourced from {[}41{]} with data available up to the 1st of August. All
model fitting was done using Markov-chain Monte Carlo (MCMC) in stan
{[}42{]} with each location and forecast target being fitted separately.

@@

For the main analysis we focused mostly on two week ahead forecasts, as
COVID-19 forecasts, especially for cases, were in the past found to have
poor predictive performance beyond this horizon {[}11{]}. Forecasts for
cases were scored using the full period from October 2020 until March
2021. To ensure comparability between models, all death forecasts were
scored using only the period from December 14th on, where all models
including the convolution model were available. To ensure robustness of
our results we conducted a sensitivity analysis where all forecasts
(including cases) were scored only over the later period for which all
forecasts were available (see \nameref{fig:agg-performance-all-late} and
\nameref{tab:score-table-late-2} and \nameref{tab:score-table-late-4}s).
Results remained broadly unchanged.

Forecasts were analysed using the following scoring metrics: The
weighted interval score (WIS) {[}43{]}, the absolute error, relative
bias, and empirical coverage of the 50\% and 90\% prediction intervals.
The WIS is a proper scoring rule {[}44{]}, meaning that in expectation
the score is optimised by reporting a predictive distribution that is
identical to the true data-generating distribution. Forecasters are
therefore incentivised to report their true belief about the future. The
WIS can be understood as a generalisation of the absolute error to
quantile-based forecasts (also meaning that smaller values are better)
and can be decomposed into three separate penalties: forecast spread
(i.e.~uncertainty of forecasts), over-prediction and under-prediction.
While the over- and under-prediction components of the WIS capture the
amount of over-prediction and under-prediction in absolute terms, we
also look at a relative tendency to make biased forecasts. The bias
metric {[}5{]} we use captures how much probability mass of the forecast
was above or below the true value (mapped to values between -1 and 1)
and therefore represents a general tendency to over- or under-predict in
relative terms. A value of -1 implies that all quantiles of the
predictive distribution are below the observed value and a value of 1
that all quantiles are above the observed value. Empirical coverage is
the percentage of observed values that fall inside a given prediction
interval (e.g.~how many observed values fall inside all 50\% prediction
intervals). Scoring metrics are explained in more detail in
\nameref{tab:scoring-metrics}. All scores were calculated using the
\texttt{scoringutils} R package {[}45{]}.

@@

To varying degrees, all forecasts exhibited trend-following behaviour
and were rarely able to predict a change in trend before it had
happened. For example, all forecasts failed to predict the change in
trend from increase to decrease that happened in November in Germany and
severely overshot reported cases (Fig \ref{fig:forecasts-and-truth}A).
This was most striking for the renewal model, which extrapolated
unconstrained exponential growth based on the recent past of
observations. The Hub ensemble and the crowd forecast, which had both
been under-predicting throughout October, also failed to predict the
change in trend after cases peaked, but less severely so. Human
forecasters, possibly aware of the semi-lockdown announced on November
2nd 2020 {[}46{]} and the change in the testing regime (with stricter
test criteria) on November 11th 2020 {[}32{]}, were fastest to adapt to
the new trend, and the Hub ensemble slowest. In December, cases rose
again in Germany, with all models under-predicting this growth to
varying extents. As in October, the renewal model captured the phase of
exponential growth in cases slightly better than other approaches, but
again overshot when reported case numbers fell over Christmas. The large
variance in predictions in January in Germany (severe under-prediction
followed by severe over-prediction) may in part be caused by the fact
that the renewal model operated on daily data and therefore was
susceptible to fluctuations in daily reporting around Christmas that
would not have influenced on weekly reporting. Similar trends in
performance were evident in Poland, with the crowd forecast quickest at
adapting to the change in trend in November. In general, there were
fewer large outlier forecasts in Poland and in particular the renewal
model performed more in line with other forecasts there.

@@

Past studies have evaluated the performance of model-based forecasting
approaches as well as human experts and non-experts in various contexts.
However, most of these studies either focused only on the evaluation of
(expert-tuned) model-based approaches {[}e.g. 8,9,10{]}, or exclusively
on human forecasts {[}15,16,20,21{]}. In contrast, we directly compared
human and model-based forecasts. This is similar to the approach taken
by {[}7{]}, but extends it in several ways. While Farrow et al.~only
asked for point predictions and constructed a predictive distribution
from these, we asked participants to provide a full predictive
distribution, allowing us to compare human forecasts and models without
any further assumptions, as well as to analyse how humans quantified
their uncertainty. In addition, we compared crowd forecasts to two
semi-mechanistic models informed by basic epidemiological knowledge of
COVID-19, allowing us to assess not only relative performance but also
to analyse qualitative differences between human judgement and
model-based insight. In terms of interpretability of the results, exact
knowledge of our two models, as well as focus on a limited set of
targets and locations was a major advantage of our study compared to
larger studies conducted by the Forecast Hubs {[}8--11,13{]}.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-johanssonOpenChallengeAdvance2019}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{Johansson MA, Apfeldorf KM, Dobson S, Devita J, Buczak
AL, Baugher B, et al. An open challenge to advance probabilistic
forecasting for dengue epidemics. PNAS. 2019;116: 24268--24274.
doi:\href{https://doi.org/10.1073/pnas.1909865116}{10.1073/pnas.1909865116}}

\leavevmode\vadjust pre{\hypertarget{ref-yamanaSuperensembleForecastsDengue2016}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{Yamana TK, Kandula S, Shaman J. Superensemble forecasts
of dengue outbreaks. Journal of The Royal Society Interface. 2016;13:
20160410.
doi:\href{https://doi.org/10.1098/rsif.2016.0410}{10.1098/rsif.2016.0410}}

\leavevmode\vadjust pre{\hypertarget{ref-colon-gonzalezProbabilisticSeasonalDengue2021}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{Colón-González FJ, Bastos LS, Hofmann B, Hopkin A,
Harpham Q, Crocker T, et al. Probabilistic seasonal dengue forecasting
in {Vietnam}: {A} modelling study using superensembles. PLOS Medicine.
2021;18: e1003542.
doi:\href{https://doi.org/10.1371/journal.pmed.1003542}{10.1371/journal.pmed.1003542}}

\leavevmode\vadjust pre{\hypertarget{ref-viboudRAPIDDEbolaForecasting2018}{}}%
\CSLLeftMargin{4. }%
\CSLRightInline{Viboud C, Sun K, Gaffey R, Ajelli M, Fumanelli L, Merler
S, et al. The {RAPIDD} ebola forecasting challenge: {Synthesis} and
lessons learnt. Epidemics. 2018;22: 13--21.
doi:\href{https://doi.org/10.1016/j.epidem.2017.08.002}{10.1016/j.epidem.2017.08.002}}

\leavevmode\vadjust pre{\hypertarget{ref-funkAssessingPerformanceRealtime2019}{}}%
\CSLLeftMargin{5. }%
\CSLRightInline{Funk S, Camacho A, Kucharski AJ, Lowe R, Eggo RM,
Edmunds WJ. Assessing the performance of real-time epidemic forecasts:
{A} case study of {Ebola} in the {Western Area} region of {Sierra
Leone}, 2014-15. PLOS Computational Biology. 2019;15: e1006785.
doi:\href{https://doi.org/10.1371/journal.pcbi.1006785}{10.1371/journal.pcbi.1006785}}

\leavevmode\vadjust pre{\hypertarget{ref-delvalleSummaryResults201420152018}{}}%
\CSLLeftMargin{6. }%
\CSLRightInline{Del Valle SY, McMahon BH, Asher J, Hatchett R, Lega JC,
Brown HE, et al. Summary results of the 2014-2015 {DARPA Chikungunya}
challenge. BMC Infectious Diseases. 2018;18: 245.
doi:\href{https://doi.org/10.1186/s12879-018-3124-7}{10.1186/s12879-018-3124-7}}

\leavevmode\vadjust pre{\hypertarget{ref-farrowHumanJudgmentApproach2017}{}}%
\CSLLeftMargin{7. }%
\CSLRightInline{Farrow DC, Brooks LC, Hyun S, Tibshirani RJ, Burke DS,
Rosenfeld R. A human judgment approach to epidemiological forecasting.
PLOS Computational Biology. 2017;13: e1005248.
doi:\href{https://doi.org/10.1371/journal.pcbi.1005248}{10.1371/journal.pcbi.1005248}}

\leavevmode\vadjust pre{\hypertarget{ref-funkShorttermForecastsInform2020}{}}%
\CSLLeftMargin{8. }%
\CSLRightInline{Funk S, Abbott S, Atkins BD, Baguelin M, Baillie JK,
Birrell P, et al. Short-term forecasts to inform the response to the
{Covid-19} epidemic in the {UK}. medRxiv. 2020; 2020.11.11.20220962.
doi:\href{https://doi.org/10.1101/2020.11.11.20220962}{10.1101/2020.11.11.20220962}}
